---
title: The Quiz (Part 2)
author: Martin Rydén
date: '2020-06-27'
slug: the-quiz-part-2
categories: []
tags:
  - how-to
  - R
  - programming
  - data-mining
  - statistics
  - web-scraping
  - wikipedia
description: ''
featured_image: ''
draft: true
---

Now that we have gathered our knowledgebase, it is time to quiz ourselves on our knowledge.

```{r, message = F}
library(tidyverse)  # For data transformation / reshaping
library(data.table) # For data transformation / reshaping
library(rvest)      # For web scraping

wikitable <- fread("wikitable.tsv")
```

```{r}
# Function to clean sentences by removing some non-alphanumeric characters and removing if it's a "common" word
clean <- function(dat){
	dat %>%
		tolower() %>%
		strsplit(split = " ") %>%
		unlist() %>% 
		gsub("[[:punct:]]", "", .) %>% 
		.[nchar(.) > 0]
}
```

```{r}

# Total unique words across all protein function descriptions
words <- paste(wikitable$protein_function, collapse = " ") %>% 
	clean()

# How many unique words are there?
n <- length(words)

quiz <- function() {
	# Select a random protein
	random_protein <- wikitable[round(runif(1, 0, .N))]
	
	# The background data: the words in the current protein's function summary
	bg <- random_protein[,"protein_function"] %>%
		clean()
	
	# Number of words in bg
	m <- length(bg)
	
	# Prompt question to player
	response <- readline(prompt = paste0("What do you know about ", random_protein, "?\t")) %>% 
		clean()
	
	# Number of words in response which are also in bg
	q <- length(response[response %in% bg])
	
	# Total number of words in response
	k <- length(response)
	
	# What is the probability of randomly selecting q words, given:
	# - q, the number of words in response which are also in m
	# - m, the number of words in the protein's function summary
	# - n, the total number of words in all proteins' function summaries
	# - k, the number of words in response
	score <- round(phyper(q - 1, m, n - m, k, lower.tail = T), 4) * 100
	
	# Rate score
	if(score > 50){
		return(paste("Good job! Your score:", score))
	} else {
		return(paste("Work harder! Your score:", score))
	}
}

```

It works! But - we can cheat by simply typing words that are common in the English language:

```{r, eval = F}
quiz()
# > What do you know about Dermatopontin?
# > it do be like as if in a cell
# > "Good job! Your score: 99.96"
```

:I

Back to Wikipedia!

https://en.wikipedia.org/wiki/Most_common_words_in_English


<iframe width="800" height="650" scrolling="yes" frameborder="yes"  src="https://en.wikipedia.org/wiki/Most_common_words_in_English#100_most_common_words" > </iframe>


```{r}
wurl <- "https://en.wikipedia.org/wiki/Most_common_words_in_English"

# Get table
wtable <- wurl %>%
	read_html %>%
	html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[1]') %>%
	html_table()

# Get words
common <- wtable[[1]]$Word

```

And let's update our 'clean' function

```{r}
# Also remove "common" words
clean <- function(dat){
	dat %>%
		tolower() %>%
		strsplit(split = " ") %>%
		unlist() %>%
		gsub("[[:punct:]]", "", .) %>% 
		unique() %>% 
		.[nchar(.) > 0] %>% 
		.[!. %in% common]
}
```

Okay, we are ready for the second attempt.

```{r}

# Total unique words across all protein function descriptions
words <- paste(wikitable$protein_function, collapse = " ") %>% 
	clean()

# How many unique words are there?
n <- length(words)

# Median number of words across function summaries
median_words <- wikitable$protein_function %>% 
  tolower() %>% 
  strsplit(split = " ") %>% 
  lengths() %>% 
  median() %>% 
  round()

# Slightly more sophisticated version of the quiz function
quiz <- function(protein = "random", reveal = T) {
	
	if(protein == "random") {
		# Select a random protein from our knowledgebase
		protein <- wikitable[round(runif(1, 0, .N))]
	} else {
		# Capitilize word and try to find in dataframe
		protein <- tolower(protein)
		substr(protein, 1, 1) <- toupper(substr(protein, 1, 1))
		protein <- wikitable[title == protein]
		if(!protein$title %in% wikitable$title) {
			# Throw error if name does not exist in df
			stop("Could not find protein name")
		}
	}
	
	# The background data: the words in the current protein's function summary
	bg <- protein[,"protein_function"] %>%
		clean()

	# Number of words in bg
	m <- length(bg)
	
	# Prompt question to player
	response <- readline(prompt = paste0("What do you know about ", protein, "?\t")) %>% 
		clean()
	
	# Number of words in response which are also in bg
	q <- length(response[response %in% bg])
	
	# Total number of words in response
	k <- length(response)
	
	# What is the probability of randomly selecting q words, given:
	# - q, the number of words in response which are also in m
	# - m, the number of words in the protein's function summary
	# - n, the total number of words in all proteins' function summaries
	# - k, the number of words in response
	score <- phyper(q-1, m, n - m, k, lower.tail = F)
	
	successes = length(response[response %in% bg])
	failures = length(response[!response %in% bg])
	
	
	# there are m correct words in a total of n words (n - m wrong)
	# 
	
	
	# Given 20 words to describe the protein,
	# what is the probability of having 2 correct guesses
	# considering a total 
	
	# The probability of out of 656 words (not accounting for -m)
	
	
	
	# lower.tail = T is the same as writing 1 - phyper(...), hence we get the probability that q words are NOT randomly selected
	
	# Note: The proper usage of phyper is with arguments: q-1, m, n-m, k
	# However, by using k we give an unproportionally high score if the player is lazy and gives a short answer.
	# Instead, we'll use the median number of words across all protein function descriptions
	# as the expected length of the response.
	
	# Also reveal answer?
	answer <- ""
	if(reveal) answer <- paste0("Answer: ", protein$protein_function)
	
	# Rate score
	if(score > 50){
		print(paste("Good job! Your score:", score))
		print(answer)
	} else {
		print(paste("Work harder! Your score:", score))
		print(answer)
	}
}

```

Let's try the same protein and give the same response as previously...

```{r, eval = F}
quiz(protein = "dermatopontin")
# > What do you know about Dermatopontin?
# > it do be like as if in a cell
# > "Work harder! Your score: 29.42"

```

We get some points for knowing 'cell' but it's still an F. Let's try again with a better response.

```{r, eval = F}
quiz(protein = "dermatopontin")
# > What do you know about Dermatopontin?
# > enhances tgfb1 activity and accelerates collagen formation
# > "Good job! Your score: 99.94"

```



### Conclusion - Part 2

We cleaned up the output from Part 1 and used it as content for a small game. A quiz, that utilized a probability distribution to determine if the player's description of a protein was better or worse than a set of randomly drawn words from the population.

In Part 3 we will attempt to generalize the concept, and create *the omni-quiz*.